---
title: "Problem Set 3"
author: "Peter Shoemaker"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(rdd)
library(reshape2)
library(Synth)
library(SCtools)
knitr::opts_chunk$set(echo = TRUE)
cat("\014") 
rm(list=ls())


setwd("C:\\Users\\Peter Shoemaker\\OneDrive\\Desktop\\Grad School\\Applied Economics & Modeling")

data = read.csv("STRT601-ProblemSet3-DataSet-adh2010-1.csv")
fips = read.csv("fips.csv")

fips <- dplyr::select(fips, c("stname", "st")) %>%
        dplyr::rename("stfips" = "st")
data <- dplyr::left_join(data, fips, by = "stfips")
```


```{r Q1}
#1. Plot per capita purchases over time, both for California and for a national average. What do
#you observe? Would a differences-in-differences approach be valid for recovering the causal
#effect of California's tobacco control program? Explain.

california <- dplyr:: filter(data, stfips == 6)

ca_plot <- ggplot(california, aes(x = year, y = cigsales)) +
  geom_point() +
  geom_line()
          
ca_plot

tobacco_avg = data %>% group_by(year)
over_time = tobacco_avg %>% summarise(
  average = mean(cigsales)
)

over_time$california_average = california$cigsales
over_time = rename(over_time, national_average = average)

mdf <- melt(over_time,id.vars="year")

final_plot <- ggplot(mdf, aes(x = year, y = value, colour = variable)) +
  geom_point()
final_plot


```
We observe that California cigarette sales fall much faster over time than the rest of the country.
A differences in differences approach would not be valid for recovering the causal effect because the
parallel worlds assumption does not hold for the entirety of the time period.

Question 2:
Use the Synth package in R to construct a synthetic control for California. Following Abadie
et al (2010), use the following characteristics: average price over 1980-1988, sales in 1980,
and sales in 1988. Match sales over the period 1980-1988. The functions dataprep and synth
are useful for this question.
```{r Q2}

df2 = data %>%
    dplyr:: filter(year >= 1980 & year <= 1988)


dataprep.out = dataprep(foo = data,
                        predictors = c("price"),
                        predictors.op = "mean",
                        time.predictors.prior = 1970:1979,
                        special.predictors = list(
                          list("price", 1980:1988, "mean"),
                          list("cigsales", 1980, "mean"),
                          list("cigsales", 1988, "mean")),
                        
                        dependent = "cigsales",
                        unit.variable = "stfips",
                        unit.names.variable = "stname",
                        time.variable = "year",
                        treatment.identifier = 6,
                        controls.identifier = c(1,8:10,13,16:23,27:33,35,37:40,42,44,45),
                        time.optimize.ssr = 1980:1988,
                        time.plot = 1970:1997
                        )
synth.out <- synth(data.prep.obj = dataprep.out)
gaps <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)
```

```{r Q2_plot1}
path.plot(synth.res = synth.out, dataprep.res = dataprep.out,
          Ylab = "cigsgales", Xlab = "year",
          Ylim = c(0, 200), Legend = c("California",
                                      "Synthetic California"), 
          Legend.position = "bottomright")
```

```{r Q2_plot2}
gaps.plot(dataprep.res = dataprep.out,synth.res = synth.out)


```

Question 3: Provide a table with theses characteristics for (1) California, (2) Synthetic
California, and (3) the national average. Does the synthetic control or the national average
more closely resemble California? How much weight is placed on each characteristic? And which
states receive the greatest weight in the synthetic control?
```{r Q3}
synth.tables <- synth.tab(
  dataprep.res = dataprep.out,
  synth.res = synth.out
)
print(synth.tables)

```
The synthetic control more closely resembles California than the national average.
The most weight is placed on the following states: New Mexico, North Dakota, and Minnesota.

```{r Q4}

```

```{r Q5}

```

```{r }

```