---
title: "PS3_Ngoh"
author: "Martin Ngoh"
date: "9/28/2021"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(rdd) 
library(gridExtra)
library(ivreg)
library(lfe)
library(scales)
library(jtools )
library(Synth)
library(ggplot2)
library(SCtools)

knitr::opts_chunk$set(echo = TRUE)
```

# Setup Data 
```{r, echo = FALSE}
cig <- read.csv("C:/Users/Martin/OneDrive/Desktop/ps3/tobacco_data.csv")

cdata <- read.csv("C:/Users/Martin/OneDrive/Desktop/ps3/tobacco_data.csv")

```


# 1.) Plot per capita purchases over time, both for California and for a national average. What do you observe? Would a differences-in-differences approach be valid for recovering the causal effect of California’s tobacco control program? Explain
```{r, echo = FALSE}
nat_avg = cig%>%
  group_by(year)%>%
  summarise(nat_AVG = mean(cigsales, na.rm = TRUE))

cig_cal = cig%>%
  filter(stfips == 6)


plot(cig_cal$year, cig_cal$cigsales, col = "blue", type = "l", 
     main = "Parallel Trends Graph 
      California VS National Average", 
     ylab = "Cig Sales", xlab = "California = Blue    National Average = Red",
     ylim = c(50,147))
par(new = TRUE)
plot(nat_avg$year , nat_avg$nat_AVG, col = "red", type = "l", xlab = "", ylab = "",
     ylim = c(50,147))


## No because the Trend in sales in California is decreasing or staying constant as the countries national average looks like it is growing ## 

```

# 2.) Use the Synth package in R to construct a synthetic control for California. Following Abadie et al (2010), use the following characteristics: average price over 1980-1988, sales in 1980, and sales in 1988. Match sales over the period 1980-1988. The functions dataprep and synth are useful for this question.
```{r, echo = FALSE}
countries <- c(1,8:10,13,16:23,27:29,30:33,35,37:39,40,42,44,45) 

dataprep.out = 
  dataprep(
    foo = cig,
    unit.variable = "stfips",
    time.variable = "year",
    treatment.identifier = 6,
    controls.identifier = countries,
    dependent = "cigsales",
    predictors = c("stfips", "price"),
    predictors.op = "mean",
    time.predictors.prior = 1980:1988,
    special.predictors = list(
      list("cigsales", c(1980:1988), "mean"),
      list("cigsales", 1980, "sum"),
      list("cigsales", 1988, "sum")
    ),
    time.optimize.ssr = 1988,
    time.plot = 1980:1997)

synth.out = synth(dataprep.out)


```

# 3.) Provide a table with these characteristics for (1) California, (2) Synthetic California, and (3) the national average. Does the synthetic control or the national average more closely resemble California? How much weight is placed on each characteristic? And which states receive the greatest weight in the synthetic control? The function synth.tab is useful for this question.
```{r, echo = FALSE}

synth.tables <- synth.tab(
  dataprep.res = dataprep.out,
  synth.res = synth.out
)
print(synth.tables)

```

# 4.) Plot cigarette purchases over time for California and Synthetic California. Include a vertical line in 1989, marking the adoption of the tobacco control program. Also plot the gap between theses lines over time. What do you conclude about the treatment effect? The functions path.plot and gaps.plot are useful for this questions.

```{r, echo = FALSE}
path.plot(synth.res = synth.out, dataprep.res = dataprep.out,
          Ylab = " Sales ", Xlab = " Year ", Ylim = c(60,150))
abline(v = 1989, col = "red")

gaps.plot(synth.out, dataprep.out)
abline(v = 1989, col="red")


```

#5.) Assess the statistical significance of the treatment effect.
```{r, echo = FALSE}

scg_Y <- dataprep.out$Y0plot%*% synth.out$solution.w
gaps <- dataprep.out$Y1plot -scg_Y

synth.tables <- synth.tab(dataprep.res = dataprep.out, synth.res = synth.out)
synth.tables

placebos <- generate.placebos(dataprep.out, synth.out)
plot_placebos(placebos, discord.extreme = FALSE)

```
